---
title: "Final Exam"
author: "Isaiah Morgan"
date: "April 12, 2018"
output: html_document
---

This is a take-home exam.  The exam will be due at 9:30 am on April 17, 2018.  You may email me your exam.  Please
do your work in a .Rmd file so that I can see your code.  Please hand in both your .Rmd file and either a .html
file or a .pdf file.  Please name your files with your last name first, followed by your first initial, followed
by the class, so my files would be named:
'FellinghamGFinal330W2018.Rmd' and 'FellinghamGFinal330W2018.pdf' or 'FellinghamGFinal330W2018.html'.  

As always, I expect your exam to be your own work.  You may use any notes and any help files available
for the programs, but searching for similar code on the web would not be appropriate.  You should not
ask for help from any person currently alive in the mortal state.

1. While the job ‘Actuary’ is frequently a top-rated job, it is incredibly difficult
to pass the exams to become an Associate and Fellow. One of the entry-level
exams is Exam P (on probability), and a small dataset of self-reported passing/not-
passing with university GPA for students who prepared for Exam P is found in *ExamP.csv*.  Read in the data, and fit a logistic model for the log(odds of passing) = $\beta_0 + \beta_1$GPA.   Print out a summary of the model.
What are $\hat\beta_0$ and $\hat\beta_1$?
```{r, echo=FALSE}
setwd('C:/Users/imoe9/Documents/School Work/STAT PROG/R Files/STAT 330/Final')
library(alr3)
library(MASS)
```
```{r}
examp <- read.csv('ExamP.csv')
attach(examp)

model <- glm(ExamP ~ GPA, family = 'binomial')
coef(model)
```



2.  The summary of the model prints out something called Wald statistics.  What is the value of the
Wald statistic to test $H_0:\beta_1 = 0$.  What is the p-value?
```{r}
summary(model)
```
The Wald statistic of the test $H_0:\beta_1 = 0$ is 2.133 with a p-value of 0.0329.

3.  Would you reject the null hypothesis at the $\alpha = 0.05$ level based on the results of number 2?

At the $\alpha = 0.05$ level I would reject the null hypothesis as $\alpha > p$

4.   There is another test of that same null hypothesis based on numbers found on the summary that is
based on the likelihood of this model and a related model.  What is the value of this test statistic?
```{r}
model$null.deviance - model$deviance
```


5.  What would the p-value be for this test statistic?
```{r}
1 - pchisq(model$null.deviance - model$deviance, 1)
```

6.  Would you reject the null hypothesis of number 2 at the $\alpha = 0.05$ based on the p-value from number 5?

Yes, I would reject the null hypothesis based on this p-value at the .05 significance level.

7.  One of the issues that we need to be concerned about is the distribution of GPA.  Produce a 
qq plot of GPA.  
```{r}
qqnorm(GPA)
qqline(GPA)
```


8.  Is there a reason to be concerned for the model in this case?

Yes, given this qq plot we should be concerned that our tests are invalid because the predictor, GPA, is not normally distributed.

9.  Produce a density plot of GPA.
```{r}
plot(density(GPA))
```


10.  What seems to be the issue with the data when you look at the density plot?
 
From the density plot it is very clear that GPA is non-normally distributed. GPA is significantly left skewed

11.  Using the powerTransform function, what transform would you suggest?
```{r}
powerTransform(GPA)
```

The power transform suggests taking GPA to the 3.5 power


12.  Try tranforming GPA by raising it to the $4^{th}$ power.  Now produce a qq plot.
```{r}
tGPA <- (GPA)^4
qqnorm(tGPA)
qqline(tGPA)
```

13.  There still seems to be a problem with normality, but the result is more symmetric and seems to have thin
tails, so lets try another fit of the log(odds of passing) as a function of GPA^4.  What are the $\hat\beta$'s
now?
```{r}
model2 <- glm(ExamP ~ tGPA, family = 'binomial')
coef(model2)
```


14.  According to the Wald statistic, is the slope of GPA^4 significantly different from 0
at the $\alpha = 0.05$ level?  What is the p-value?
```{r}
summary(model2)
```

The Wald statistic is 2.150 for the slope of tGPA and the p-value is 0.0315 meaning that the slope of GPA^4 is
significantly greater than 0

15.  According to the model in 13, what is the point estimate for the log(odds of passing) if the GPA is 3.1?
```{r}
bhat <- coef(model2)
est <- c(1,(3.1)^4)%*%bhat
est
```

16.  What is the point estimate for the probability of passing with a GPA of 3.1?
```{r}
exp(est)/(1+exp(est))
```

17.  Using the model in 13, what does the GPA need to be for the probability of passing to be at least 0.5? (Work to 
2 decimal places.)
```{r}
est1 <- c(1,(3.9311)^4)%*%bhat
exp(est1)/(1+exp(est1))
```
So to be totally honest I played around with numbers until I got a GPA that would work. A GPA of 3.9311 will correlate
to a .50 passing rate on exam P according to the model from question 13.


18.  Produce a plot of the probability of passing as a function of GPA using the model in problem 13.
```{r}
X <- cbind(1,13:241) 
yhat1 <- exp(X%*%bhat)/(1+exp(X%*%bhat))
plot((seq(13:241))^(1/4), yhat1, type = 'l')

```

The data in the file called *bird1.dat* have to do with the status of various bird
populations.  We are going to attempt to build a model to estimate the variable 'Status'
from the other variables in the data set (excluding 'Species').  There are 67 observations
in the data set.  We are going to build a model using 50 of the observations, and test the
model on the remaining 17.  Use the following code to make a training data set and a
test data set.  

```{r}
bird <- read.table('bird1.dat',header=TRUE)
set.seed(23)
aa <- 1:67
train <- sample(aa,50)
trainset <- bird[train,]
testset <- bird[-train,]
```

19.  Build a model on the training set using the following variables: Mass, Range, Migr, Insect,
Diet, Wood, Water, Indiv.  Print a summary of the model.  What is the residual deviance of
the model?
```{r}
birdmod <- glm(Status ~ Mass + Range + Migr + Insect + Diet + Wood + Water + Indiv, 
               data = trainset, family = 'binomial')
summary(model)
```
The residual deviance of the model is 45.653.

20.  Now perform a stepwise procedure on the training set using the model above as the full model and the model
with only the intercept as the minimal model.  Use AIC as the criterion and step both directions.  What terms are in the best model?
```{r}
birdmin <- glm(Status ~ 1, data = trainset, family = 'binomial')
step(birdmod, scope=list(upper=birdmod,lower=birdmin), direction = 'both', data = trainset) 
```
The terms of the best model from this process are Mass, Diet, Wood, and Indiv.


21. Print a summary of the best model.  What is the residual deviance of this model?
```{r}
birdm1 <- glm(Status ~ Mass + Diet + Wood + Indiv, family = "binomial", data = trainset)
summary(birdm1)
```
The residual deviance of this model is 18.428.

22.  Now do the same stepwise procedure on the training set using a larger penalty.
Use k=7 in this run.
Now what terms are in the
best model?
```{r}
step(birdmod, scope=list(upper=birdmod,lower=birdmin), direction = 'both', data = trainset, k=7) 
```
This method produces a model including Mass, Diet, and Indiv.

23.  Print a summary of the best model using k=7.  What is the residual deviance of
this model?
```{r}
birdm2 <- glm(Status ~ Mass + Diet + Indiv, family = "binomial", data = trainset)
summary(birdm2)
```
The residual deviance is 24.356

24. What is the p-value for the coefficient of Wood in the AIC model?

The p-value for Wood in the AIC model is 0.04661

25.  What is the p-value for the likelihood ratio test comparing the model using AIC
with the model using k=7?
```{r}
anova(birdm2, birdm1)
1 - pchisq(birdm2$deviance - birdm1$deviance, 1)
```

26.  We will call the model using AIC the AIC model and the model using k=7 the 
k7 model. Now predict the 17 observations in the test set using the AIC model.
What is the error sum of squares for the predicted values?
```{r}
birdt1 <- glm(Status ~ Mass + Diet + Wood + Indiv, family = "binomial", data = testset)
X1test <- model.matrix(birdt1)
beta1 <- coef(birdt1)
p1 <- exp(X1test%*%beta1) / (1+exp(X1test%*%beta1)) 
m1SSE <- sum((testset$Status - p1)^2, na.rm = TRUE)
m1SSE
```


27.  What is the median absolute deviation for the AIC model?
```{r}
m1MAD <- median(abs(testset$Status - p1), na.rm = TRUE)
m1MAD
```

28.  Now predict the 17 observations in the test set using the k7 model.
What is the error sum of squares for the predicted values?
```{r}
birdt2 <- glm(Status ~ Mass + Diet + Indiv, family = "binomial", data = testset)
X2test <- model.matrix(birdt2)
beta2 <- coef(birdt2)
p2 <- exp(X2test%*%beta2) / (1+exp(X2test%*%beta2)) 
m2SSE <- sum((testset$Status - p2)^2, na.rm = TRUE)
m2SSE
```

29.  What is the median absolute deviation for the k7 model?
```{r}
m2MAD <- median(abs(testset$Status - p2), na.rm = TRUE)
m2MAD
```

30.  Based on all the output we have looked at, which model would you use to
predict Status in the future?

I would rely more on the model produced with k = 7 because both its SSE and MAD are lower than
the AIC model and is more parsimonious as well.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```