---
title: "Shift Report"
author: "Isaiah M."
date: "March 21, 2018"
output: pdf_document
---
```{r, include=FALSE}
setwd("C:/Users/imoe9/Documents/School Work/STAT PROG/R Files/Work")
shifts <- read.csv('SHIFTS.csv')
attach(shifts)
par(mfrow=c(1,3))
```

```{r, echo = FALSE, include = FALSE}
AVGPlghr <- rowMeans(as.matrix(shifts[,2:11]), na.rm = TRUE)
AVGChr <- rowMeans(as.matrix(shifts[,12:21]), na.rm = TRUE)
AVGAmt <- rowMeans(as.matrix(shifts[,12:21]), na.rm = TRUE)

plot(c(1:17), AVGPlghr, xlab = 'Shift', ylab = 'Pledges Per Hour', main = 'Shift Performance: Pledges per Hour',
     type = 'o' )
plot(c(1:17), AVGChr, xlab = 'Shift', ylab = 'Charge Per Hour', main = 'Shift Performance: Charge per Hour',
     type = 'o' )
plot(c(1:17), AVGAmt, xlab = 'Shift', ylab = 'Average Amount Per Hour', main = 'Shift Performance: Average Amount',
     type = 'o' )

```


```{r, echo=FALSE, include = FALSE}
#Red Dashlined shows the performance of Tuesday Early shifts (Shift 5) over a ten week period from the begining of this semester. Overall it performs below average which becomes more important as we moved into our A donor campaign.

#Pledge Per Hour
plot(NULL,NULL, xlim = c(1,10), ylim = c(0,5), ylab = 'Pledge per Hour', xlab = 'Week', main = 'Pledge Per Hour for Jan-Mar')
for (i in 1:17) {
  lines(c(1:10), shifts[i, 2:11])
}
lines(c(1:10), shifts[5,2:11],col='red', lwd = 5, lty = 6)

#Charge per Hour
plot(NULL,NULL, xlim = c(1,10), ylim = c(0,3), ylab = 'Charge per Hour', xlab = 'Week', main = 'Charge Per Hour for Jan-Mar')
for (i in 1:17) {
  lines(c(1:10), shifts[i, 12:21])
}
lines(c(1:10), shifts[5,12:21],col='red', lwd = 5, lty = 6)

#Amount per Hour
plot(NULL,NULL, xlim = c(1,10), ylim = c(0,500), ylab = 'Amount per Hour', xlab = 'Week', main = 'Amount Per Hour for Jan-Mar')
for (i in 1:17) {
  lines(c(1:10), shifts[i, 22:31])
}
lines(c(1:10), shifts[5,22:31],col='red', lwd = 5, lty = 6)
```

The following plots only look at the first 6 weeks of calling when we were calling only prospective donors. In order to include the
other weeks I need more information about which donor groups we were calling to perform a statistical analysis that accurately reflects the truth. According to the analysis of variance tables I have shown below none of the shifts performed
significantly different from the others while we were calling prospectives. 

This is further apparent if we look at boxplots of each indicator by shift which fails to show any statistically
significant differences between shifts.


```{r , echo = FALSE}
#Model and Boxplots
shiftsd <- as.data.frame(shifts)
shiftplg <- as.numeric(t(shiftsd[,2:7]))
shiftchr <- as.numeric(t(shiftsd[,12:17]))
shiftamt <- as.numeric(t(shiftsd[,22:27]))
shiftvector <- c(rep('SE',6), rep('S1',6),rep('S2',6), rep('S3',6),rep('TE',6), rep('T1',6),
                 rep('T2',6), rep('T3',6),rep('W1',6), rep('W2',6),rep('W3',6), rep('TH1',6),
                 rep('TH2',6),rep('TH3',6),rep('FE',6), rep('F1',6),rep('F2',6))
shiftT <- cbind(shiftplg,shiftchr,shiftamt,shiftvector)
shiftT <- as.data.frame(shiftT)
shiftT$shiftplg <- as.numeric(as.character(shiftT$shiftplg))
shiftT$shiftchr <- as.numeric(as.character(shiftT$shiftchr))
shiftT$shiftamt <- as.numeric(as.character(shiftT$shiftamt))

fitplg <- lm(shiftT[,1] ~ shiftT[,4])
fitchr <- lm(shiftT[,2] ~ shiftT[,4])
fitamt <- lm(shiftT[,3] ~ shiftT[,4])

par(mfrow=c(2,2))
plot(fitplg)
plot(fitchr)
plot(fitamt)

anova(fitplg)
anova(fitchr)
anova(fitamt)

par(mfrow=c(1,1))
boxplot.matrix(as.matrix(shiftsd[1:17,2:7]), use.cols = FALSE, ylab = 'Pledge per Hour', xlab = 'Shift', 
               main = 'Pledge Per Hour for Jan-Feb')
boxplot.matrix(as.matrix(shiftsd[1:17,12:17]), use.cols = FALSE, ylab = 'Charge per Hour', xlab = 'Shift', main =
              'Charge Per Hour for Jan-Feb')
boxplot.matrix(as.matrix(shiftsd[1:17,22:27]), use.cols = FALSE, ylab = 'Amount per Hour', xlab = 'Shift', 
               main = 'Amount Per Hour for Jan-Feb')  
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

